---
title: "SDS_Exercise3"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
library(FNN)
greenbuildings = read.csv("data/greenbuildings.csv")
wine = read.csv("data/wine.csv")
social_marketing = read.csv("data/social_marketing.csv")
set.seed(3)
rmse = function(y, yhat) {
  sqrt( mean( (y - yhat)^2 ) )
}
n = nrow(greenbuildings)
n_train = round(0.8*n) # round to nearest integer n_test = n ­ n_train
gb_no_green = subset(greenbuildings, green_rating == 0)
gb_no_green$green_rating[gb_no_green$green_rating == 0] <- 1
gb_green = subset(greenbuildings, green_rating == 1)
gb_green$green_rating[gb_green$green_rating == 1] <- 0


rmse_vals = do(100)*{
  
  # re-split into train and test cases with the same sample sizes
  train_cases = sample.int(n, n_train+1, replace=FALSE) 
  test_cases = setdiff(1:n, train_cases) 
  buildings_train = greenbuildings[train_cases,] 
  buildings_test = greenbuildings[test_cases,]
  
  lm1 = lm(Rent ~ renovated + size + stories + age + green_rating, data=buildings_test)
  lm2 = lm(Rent ~ (Gas_Costs + Electricity_Costs + total_dd_07)^2 + 
             (age + renovated)^2 + 
             (Gas_Costs + Electricity_Costs + green_rating)^2 +
             stories + 
             cluster_rent + 
             size, data=buildings_test)
  lm4 = lm(Rent ~ amenities + green_rating, data=buildings_test)
  lm5 = lm(Rent ~ Gas_Costs + Electricity_Costs + total_dd_07 + green_rating, data=buildings_test)
  
  # Predictions out of sample
  yhat_test1 = predict(lm1, buildings_test)
  yhat_test2 = predict(lm2, buildings_test)
  yhat_test4 = predict(lm4, buildings_test)
  yhat_test5 = predict(lm5, buildings_test)
  
  nogreen_to_green = predict(lm2, gb_no_green)
  green_to_nogreen = predict(lm2, gb_green)
  
  
  c(rmse(buildings_test$Rent, yhat_test1),
    rmse(buildings_test$Rent, yhat_test4),
    rmse(buildings_test$Rent, yhat_test5),
    rmse(buildings_test$Rent, yhat_test2),
    mean(nogreen_to_green - gb_no_green$Rent),
    mean(green_to_nogreen - gb_green$Rent)
    
  )
}







```



# Predictive Model Building

### Problem
For this first question, our goal was to create a model that would be able to predict the rent of buildings from Exercise 1. From there, we should be able to take that model, and predict the change in rent if a building's green_rating changed.

### Data and Model
We created 3 linear regression models to test various combinations of attributes to see if they were good indicators of Rent. The first (V1) was a combination of attributes related to age and size, the second (V2) was only concerned with amenities, and the last (V3) looked at gas/electricity costs and the number of heating/cool days in the region. All three of these models also used the green_rating. We decided to lump LEED and EnergyStar together for simplicity as we did not see too much difference between the two. The green_rating attribute indicates whether or not the building was green certified.

In our fourth linear regression model (V4), we combined all of these elements and introduced interactions between attributes. We included interactions between gas/electricity costs and total degree days as the number of degree days will affect the cost of gas/electricity. We included an interaction between age and renovation as a building that is old but renovated will likely be more expensive than a building that is only old and not renovated. The final interaction we had was between gas/electricity costs and green_rating, as a building will be more energy efficient if it is a 'green' building, resulting in lower gas/electricity costs. Other than those interactions, we also used stories, size, and cluster_rent as they were good/significant predictors.

We tested these models on 100 different test/training splits. For each split, we created the three linear models on a training set, predicted Rent on a given test set, and calculated RMSE between the four models' predictions and actual values for Rent. We selected different test/training sets for each split.

As for predicting the average change in Rent given a change in green_rating, we took the entire dataset, flipped the green_rating, and predicted the Rent on the entire set using V4 with all other attributes held constant. Once we had that value, we took the mean of the differences between the predicted Rent and actual Rent of every building. We did this within the test/training splits as well.

### Results
```{r green_buildings,  echo=FALSE}

boxplot(rmse_vals[1:4],xlab = "Models", ylab = "RMSE of Rent (dollars/sq ft)", main = "RMSE of Linear Regression Models")
colMeans(rmse_vals[1:4])

```
As seen above, the final model V4 outperforms the other three due to the use of interactions as well as incorporating elements from the other three models.

```{r green_buildings_2,  echo=FALSE}

boxplot(rmse_vals[5:6], xlab = "Changes to Green Rating", ylab = "RMSE of Rent (dollars/sq ft)", main = "Difference in Dollar Amounts")
colMeans(rmse_vals[5:6])

```
V5 indicates the average change in Rent when a building that was previously not green becomes green. V6 indicates the average change in Rent when a building that was previously green becomes not green. As seen above, if a building changes its green rating, the change in Rent is approximately 1 dollar according to our model. That is, rent will be \$1.08 more expensive if a building becomes green, and \$1.05 less expensive if a building loses its green status.

### Conclusion
We were able to create a model using linear regression techniques to predict a building's rent, where the average RMSE of our model was 9.34. From that model, we were able to project the change in rent given a change in a building's green_rating while all other factors constant. It appears if a building's green_rating status changes, the change in rent will be around a dollar, either more expensive if the building was previously not green or less expensive if the building was previously green.

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
